---
title: "Preparing (munging) data for analysis"
author: "Marin Grgurev"
date: "Thursday, September 11, 2014"
output: html_document
---

```{r Read raw data, include=FALSE}
if(basename(getwd()) == "reports") setwd("..")
library(data.table)
load("data/data.Rda")
data <- setDT(data)
```
Adjusting the variable class and names
This script prepare the raw data obtained  from the PG database to use in later analysis. 

The steps to prepare the tidy data set includes:

* creating of data.table (raw data is approximately ~ 1 million rows with 6 variables big dataframe)

* adjusting variable names to be inline with the Google's R Style Guide guidance regarding easier reading, sharing and verifying of R code

* concatenating name of the ecological variables with name of the statistic and removing of unused statistic column

* convert pres.abs and radius.hr variables to factors

* caching tidy dataset for later use

### Changing the name for some of the variables
```{r Setnames}
setnames(data, colnames(data)[c(2,3,4,5,6)], c("pres.abs", "radius.hr", "ecol.var", "statistic", "value"))
```

### Concatenating all variable names with statistic except lvoc variables and removing statistic variable
```{r Concatenating variables with statistic}
data[statistic != "percent", ecol.var := paste(data[statistic != "percent", ecol.var], "_", data[statistic != "percent", statistic], sep = "")]
data[,statistic := NULL]
```

### Change type for pres.abs and radius.hr variables to factors
```{r}
data[,pres.abs := as.factor(pres.abs)]
data[,radius.hr := as.factor(radius.hr)]
```

### Cache data
```{r}
#cache("data")
```

